Characterizing Signal Loss in the 21\,cm Reionization Power Spectrum: \\A Revised Study of PAPER-64

### 2018-10-08 PAPER.tex
-Removed appendices and all references to them :( -Clarified statement about sources of signal loss in the analysis pipeline pre-PS

Intuitively, the estimator must be capable of ``suppressing" or ``removing" the effects of contaminants in order to obtain an unbiased estimate of the power spectrum. By construction, the subtraction of the residual foreground and noise bias accomplishes this, removing any additive bias. However, the $\mathbf{C}^{-1}$ piece of Equation \eqref{eq:OQEQuadratic} also has the effect of suppressing residual foregrounds and noise, in both the additive bias and any contributions the residuals may have to the variance. The way in which the optimal estimator accomplishes this is illustrated with a toy model in Appendix \ref{sec:icw_appendix}. In this appendix, we show that the effect of the weighting in Equation \eqref{eq:OQEQuadratic} is to project out the modes of $\textbf{U}$ with a different covariance structure than $\mathbf{S}$ in the power spectrum estimate, and the effect of Equation \eqref{eq:OQELinear} is to subtract out the remaining bias. Similar effects for a realistic model of the EoR and foregrounds are shown in \citet{liu_tegmark2011}.

The toy model in Appendix \ref{sec:icw_appendix} also illustrates that if the covariance structure of the contaminants is sufficiently different from the desired power spectrum, then the linear bias term may be expected to be quite small, and it is only necessary to know $\C$ and $\textbf{Q}^{\alpha}$, but not $\mathbf{U}$. Since the foregrounds are expected to be strongly correlated between frequencies whereas the EoR is not, we expect different covariance structures and therefore a small linear bias. Moreover, because the linear bias is always positive and there is no multiplicative bias, the quadratic-only term will always produce an estimate which is {\it high} relative to the true value, and which can conservatively be interpreted as an upper limit. These considerations, and the difficulty of obtaining an estimate for $\mathbf{U}$, motivate the neglect of the linear bias in the rest of this analysis.

@@

As discussed in Section \ref{sec:QE}, this behavior is {\it not} expected in the case that we were to use a true $\textbf{C}^{-1}$ weighting.  Rather, we would obtain the behavior shown in the toy model in Appendix \ref{sec:icw_appendix}\footnote{Note that there the true covariance matrix is also the sum of a diagonal portion describing the signal, and a single mode describing the contaminant (similar to Figure \ref{fig:toy_sigloss12}).}, with suppression of the foreground mode resulting in a nearly unbiased estimate of the power spectrum.   The key difference is that since $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it, and this coupling leads to loss.

For the case of an eigenmode which can be safely assumed to be predominantly a foreground, its presence in the true covariance matrix will result in the desired suppression via a kind of projection (Equations \eqref{eq:IdealToyModelCovariance} and \eqref{eq:IdealToyModelEstimator}); whether or not it is strongly correlated with the the actual data vector is irrelevant.  However, in the case of an empirically estimated covariance matrix, the eigenmodes of $\widehat{\textbf{C}}_{\rm EoR}$ will both be incorrect and can be correlated with the data. If these incorrect eigenmodes are not correlated with the data, it will lead to non-minimum variance estimates but will not produce the suppression of the power spectrum amplitude as seen in the left plot of Figure \ref{fig:toy_sigloss3}. As described in Section \ref{sec:siglossmethod} and shown mathematically in Appendix \ref{sec:sigloss_appendix}, however, if $\widehat{\mathbf{C}}_{\rm EoR}$ $\textit{is}$ correlated with the data vector $\mathbf{x}$, there is a kind of projection of power in the {\it non}-foreground modes from the resulting power spectrum estimate, thus producing an estimate that is biased low.  In short, {\it if the covariance is computed from the data itself, it carries the risk of overfitting information in the data and introducing a multiplicative bias (per $k$) to estimates of the signal.}

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the EoR-dominated eigenmodes of $\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. More specifically, the coupling between the estimated covariance and the data is anti-correlated in nature (which is explained in more detail in Section \ref{sec:siglossmethod} and Appendix \ref{sec:sigloss_appendix}), which leads to loss. Mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is therefore more harmful than for FG-dominated modes, and since the lowest-valued eigenmodes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

@@

The situation here with using a finite number of time samples to estimate our covariance is analogous to a problem faced in galaxy surveys, where the non-linear covariance 
of the matter power spectrum is estimated using a large --- but finite --- number of expensive simulations. There, the limited 
number of independent simulations results in inaccuracies in estimated covariance matrices 
\citep{dodelson_schneider2013,taylor_joachimi_etal2014}, which in turn result in biases in the final parameter constraints 
\citep{hartlap_et_al2007}. In our case, the empirically estimated covariances are used for estimating the power spectrum, and 
as we discussed in the previous section (and will argue more thoroughly in Section \ref{sec:siglossmethod} and Appendix \ref{sec:sigloss_appendix}), couplings between these covariances and the data can lead to power spectrum estimates that are biased 
\emph{low}---which is precisely signal loss. In future work, it will be fruitful to investigate whether advanced techniques from the 
galaxy survey literature for estimating accurate covariance matrices can be successfully adapted for $21\,\textrm{cm}$ 
cosmology. These techniques include the imposition of sparsity priors \citep{padmanabhan_et_al2016}, the fitting of 
theoretically motivated parametric forms \citep{pearson_samushia2016}, covariance tapering \citep{paz_sanchez2015}, 
marginalization over the true covariance \citep{sellentin_heavens2016}, and shrinkage methods 
\citep{pope_szapudi2008,joachimi_2017}.

@@

Finally, we emphasize that the discussion that follows is solely focused on signal loss associated with empirical covariance weighting. As mentioned in Section \ref{sec:SiglossOverview}, there are several other ways the PAPER analysis pipeline can lead to loss, including through calibration and delay-filtering. We refer the reader to \citet{parsons_et_al2010} and \citetalias{ali_et_al2015} for discussions about these other types of signal loss, noting here that they are much less significant than the loss focused on in this paper.

@@

{\bf In general:} Another way to phrase the shortcoming of the empirical inverse covariance estimator (which is also discussed in Appendix \ref{sec:sigloss_appendix}) is that it is not properly normalized. Signal loss due to couplings between the data and its weightings arise because our unnormalized quadratic estimator from Equation \eqref{eq:qhat} ceases to be a quadratic quantity, and instead contains higher order powers of the data. However, the normalization matrix $\mathbf{M}$ is derived assuming that the unnormalized estimator is quadratic in the data. The power spectrum estimate will therefore be incorrectly normalized, which manifests as signal loss. We leave a full analytic solution for $\mathbf{M}$ for future work, since our simulations already capture the full phenomenology of signal loss and have the added benefit of being more easily generalizable in the face of non-Gaussian systematics.

@@

Going forward, we treat every $k$-value separately in order to determine an upper limit on the EoR signal per $k$. We bin our simulation outputs along the $P_{\rm in}$ axis (one bin per injection level) and, since they are well-approximated by a Gaussian distribution in our numerical results, we smooth the distribution of $\widehat{P}_{\rm out}$ values by fitting Gaussians for each bin based on its mean and variance (and normalize them). Stitching all of them together results in a 2-dimensional transfer function --- the likelihood function in Bayes' rule, namely $\mathcal{L} (  \widehat{P}_{\rm out} | P_{\rm in})$. We then have a choice for our prior, $p(P_{\rm in})$, and we choose to invoke a Jeffreys prior (\citealt{jaynes1968}) because it is a true uninformative prior. For a derivation and more details about the Jeffreys prior used in our analysis, see Appendix \ref{sec:jeffreys}.

@@

>>>>

Intuitively, the estimator must be capable of ``suppressing" or ``removing" the effects of contaminants in order to obtain an unbiased estimate of the power spectrum. By construction, the subtraction of the residual foreground and noise bias accomplishes this, removing any additive bias. However, the $\mathbf{C}^{-1}$ piece of Equation \eqref{eq:OQEQuadratic} also has the effect of suppressing residual foregrounds and noise, in both the additive bias and any contributions the residuals may have to the variance.

More specifically, the effect of the weighting in Equation \eqref{eq:OQEQuadratic} is to project out the modes of $\textbf{U}$ with a different covariance structure than $\mathbf{S}$ in the power spectrum estimate, and the effect of Equation \eqref{eq:OQELinear} is to subtract out the remaining bias. Similar effects for a realistic model of the EoR and foregrounds are shown in \citet{liu_tegmark2011}. If the covariance structure of the contaminants is sufficiently different from the desired power spectrum, then the linear bias term may be expected to be quite small, and it is only necessary to know $\C$ and $\textbf{Q}^{\alpha}$, but not $\mathbf{U}$. Since the foregrounds are expected to be strongly correlated between frequencies whereas the EoR is not, we expect different covariance structures and therefore a small linear bias. Moreover, because the linear bias is always positive and there is no multiplicative bias, the quadratic-only term will always produce an estimate which is {\it high} relative to the true value, and which can conservatively be interpreted as an upper limit. These considerations, and the difficulty of obtaining an estimate for $\mathbf{U}$, motivate the neglect of the linear bias in the rest of this analysis.

@@

As discussed in Section \ref{sec:QE}, this behavior is {\it not} expected in the case that we were to use a true $\textbf{C}^{-1}$ weighting.  Rather, we would obtain %the behavior shown in the toy model in Appendix \ref{sec:icw_appendix}\footnote{Note that there the true covariance matrix is also the sum of a diagonal portion describing the signal, and a single mode describing the contaminant (similar to Figure \ref{fig:toy_sigloss12}).}, with suppression of the foreground mode resulting in 
a nearly unbiased estimate of the power spectrum.   The key difference is that since $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it, and this coupling leads to loss.

For the case of an eigenmode which can be safely assumed to be predominantly a foreground, its presence in the true covariance matrix will result in the desired suppression via a kind of projection; whether or not it is strongly correlated with the the actual data vector is irrelevant.  However, in the case of an empirically estimated covariance matrix, the eigenmodes of $\widehat{\textbf{C}}_{\rm EoR}$ will both be incorrect and can be correlated with the data. If these incorrect eigenmodes are not correlated with the data, it will lead to non-minimum variance estimates but will not produce the suppression of the power spectrum amplitude as seen in the left plot of Figure \ref{fig:toy_sigloss3}. As described in Section \ref{sec:siglossmethod}, however, if $\widehat{\mathbf{C}}_{\rm EoR}$ $\textit{is}$ correlated with the data vector $\mathbf{x}$, there is a kind of projection of power in the {\it non}-foreground modes from the resulting power spectrum estimate, thus producing an estimate that is biased low.  In short, {\it if the covariance is computed from the data itself, it carries the risk of overfitting information in the data and introducing a multiplicative bias (per $k$) to estimates of the signal.}

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the EoR-dominated eigenmodes of $\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. More specifically, the coupling between the estimated covariance and the data is anti-correlated in nature (which is explained in more detail in Section \ref{sec:siglossmethod}), which leads to loss. Mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is therefore more harmful than for FG-dominated modes, and since the lowest-valued eigenmodes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

@@

The situation here with using a finite number of time samples to estimate our covariance is analogous to a problem faced in galaxy surveys, where the non-linear covariance 
of the matter power spectrum is estimated using a large --- but finite --- number of expensive simulations. There, the limited 
number of independent simulations results in inaccuracies in estimated covariance matrices 
\citep{dodelson_schneider2013,taylor_joachimi_etal2014}, which in turn result in biases in the final parameter constraints 
\citep{hartlap_et_al2007}. In our case, the empirically estimated covariances are used for estimating the power spectrum, and 
as we discussed in the previous section (and will argue more thoroughly in Section \ref{sec:siglossmethod}), couplings between these covariances and the data can lead to power spectrum estimates that are biased 
\emph{low}---which is precisely signal loss. In future work, it will be fruitful to investigate whether advanced techniques from the 
galaxy survey literature for estimating accurate covariance matrices can be successfully adapted for $21\,\textrm{cm}$ 
cosmology. These techniques include the imposition of sparsity priors \citep{padmanabhan_et_al2016}, the fitting of 
theoretically motivated parametric forms \citep{pearson_samushia2016}, covariance tapering \citep{paz_sanchez2015}, 
marginalization over the true covariance \citep{sellentin_heavens2016}, and shrinkage methods 
\citep{pope_szapudi2008,joachimi_2017}.

@@

Finally, we emphasize that the discussion that follows is solely focused on signal loss associated with empirical covariance weighting. As mentioned in Section \ref{sec:SiglossOverview}, there are a number of steps in our analysis pipeline which could lead to loss, including gain calibration, delay filtering, and fringe-rate filtering, which have been investigated at various levels of detail in \citet{parsons_et_al2014} and \citetalias{ali_et_al2015} but are clearly the subject of future work. Here we only focus on the most significant source of loss we have identified and note that other sources of loss may still be present (though at much smaller levels).

@@

{\bf In general:} Another way to phrase the shortcoming of the empirical inverse covariance estimator is that it is not properly normalized. Signal loss due to couplings between the data and its weightings arise because our unnormalized quadratic estimator from Equation \eqref{eq:qhat} ceases to be a quadratic quantity, and instead contains higher order powers of the data. However, the normalization matrix $\mathbf{M}$ is derived assuming that the unnormalized estimator is quadratic in the data. The power spectrum estimate will therefore be incorrectly normalized, which manifests as signal loss. We leave a full analytic solution for $\mathbf{M}$ for future work, since our simulations already capture the full phenomenology of signal loss and have the added benefit of being more easily generalizable in the face of non-Gaussian systematics.

@@

Going forward, we treat every $k$-value separately in order to determine an upper limit on the EoR signal per $k$. We bin our simulation outputs along the $P_{\rm in}$ axis (one bin per injection level) and, since they are well-approximated by a Gaussian distribution in our numerical results, we smooth the distribution of $\widehat{P}_{\rm out}$ values by fitting Gaussians for each bin based on its mean and variance (and normalize them). Stitching all of them together results in a 2-dimensional transfer function --- the likelihood function in Bayes' rule, namely $\mathcal{L} (  \widehat{P}_{\rm out} | P_{\rm in})$. We then have a choice for our prior, $p(P_{\rm in})$, and we choose to invoke a Jeffreys prior (\citealt{jaynes1968}) because it is a true uninformative prior. %For a derivation and more details about the Jeffreys prior used in our analysis, see Appendix \ref{sec:jeffreys}.

@@

\begin{comment}
